---
title: "Tme Series Analysis with R"
author: "Craig W. Slinkman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intorduction



Hello.  My name is Craig Slinkman and I am a retired professor of business 
statistics.  From `r 2017 - 1979 + 1` 
teaching business statistics in the College of Business Statistics at the
University of Texas at Arlington.  The great majority of my work is in applied 
statistical analysis.

I am an avid R user.   I use RStudio and mu IDE.  I make extensive use of  tidyverse.   
This document was produced in RStudio. 

This is an introduction of how to use the tidyverse to produce forecasting models.
It is made from a number of tutorials that can be found on the Internet.

The topic of this document is based on vignette which can be found [here](https://cran.rstudio.com/web/packages/tsibble/vignettes/intro-tsibble.html).

We start with a variant of a tibble called the tsibble:

# My version of the tsibble vignette.

## Introduction to tsibble  
Earo Wang as modified by Craig Slinkman  

The tsibble package extends the tidyverse to temporal data. Built on top of the 
tibble, a tsibble (or tbl_ts) is a data- and model-oriented object. Compared to 
the conventional time series objects in R, for example ts, zoo, and xts, the 
tsibble preserves time indices as the essential data column and makes 
heterogeneous data structures possible. Beyond the tibble-like representation, 
key comprised of single or multiple variables is introduced to uniquely 
identify observational units over time (index). The tsibble package aims at 
managing temporal data and getting analysis done in a fluent workflow.

### Contextual semantics: index and key  

tsibble() creates a **tsibble** object, and **as_tsibble()** is an S3 method to 
coerce other objects to a **tsibble**. An object that a vector/matrix 
underlies, such as **ts** and **mts**, can be automated to a **tsibble** using 
**as_tsibble()** without any specification. If it is a tibble or 
data frame,**as_tsibble()** requires a little more setup in order to declare the 
index and key variables.

##### Slinkman additional remarks

We first need to load the required packages.  Note that I use the **require()** 
instead of the **library()** function.  They both load the required packages.  
The difference is that the **required()** function returns a return code 
indicating either success or failure.  

We load the packages below

```{r load_packages}
library(tidyverse)                            # Automatically loads dplyr.
library(lubridate)                            # For dates.
library(tsibble)                              # For tidyverse time series data.
```

We now display selected columns of the **tibble**  **nycflights13::weather** 
using **dplyr**

```{r display_weather}
weather <- nycflights13::weather %>% 
  select(origin, time_hour, temp, humid, precip)
weather
```
  
The weather data included in the package nycflights13 contains the hourly meteorological records (such as temperature, humid and precipitation) over the year of 2013 at three stations (i.e. JFK, LGA and EWR) in New York City. Since the **time_hour** is the only column involving the **timestamps**, **as_tsibble()** defaults it to the index variable; alternatively, the index can be specified by the argument index = time_hour to disable the verbose message.

Except for index, a tsibble requires “key”, which defines subjects or individuals measured over time. In this example, the origin variable is the identifier, which is passed to the argument key in **as_tsibble()**. __Each observation should be uniquely identified by **index** and **key** in a valid tsibble.__ Others—temp, humid and precip—are referred to as measured variables. When creating a **tsibble**, the key will be sorted first, followed by arranging time from past to recent.

We demonstrate this here:

```{r corece_tibblr_to_tsibble}
weather_tsbl <-                             # Using `time_hour` as index 
  as_tsibble( weather,                      # variable by default.    
              key = origin )

weather_tsbl                             # Verify tsibble.                     
```



